---
title: "Covid-19 Report"
author: Ömür Özkir
date: April 22, 2020
output:
  pdf_document:
    toc: true
    toc_depth: 3
    fig_width: 7
    df_print: kable
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

\clearpage
# The dataset

The dataset used is available at [Github](https://ramikrispin.github.io/coronavirus/) under https://ramikrispin.github.io/coronavirus/ and updated daily (but might be delayed by up to a day).

The source additionally provides the data as csv file and also offers a comprehensive dashboard based on this dataset.

```{r echo=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(readr)
library(ggplot2)
library(ggthemes)
library(tidyr)
library(knitr)

# to update the dataset:
# devtools::install_github("covid19r/coronavirus", force=TRUE)
library(coronavirus)
data(coronavirus)
```


# The Last 24 Hours

## Worldwide Confirmed Cases

```{r echo=FALSE}
total <- coronavirus %>% 
  filter(type == "confirmed") %>% 
  summarise( latest_report = max(date), confirmed_cases = sum(cases))
total
```

## Total cases by country

```{r echo=FALSE}
summary_df <- coronavirus %>% group_by(Country.Region, type) %>%
  summarise(total_cases = sum(cases)) %>%
  arrange(-total_cases)

head(summary_df, 10)
```

### Total deaths per country

```{r echo=FALSE}
deaths_df <- summary_df %>% filter(type == "death") %>% select(Country.Region, total_cases)
head(deaths_df, 10)
```


```{r echo=FALSE}
newest <- coronavirus %>% 
  filter(date == max(date)) %>%
  select(country = Country.Region, type, cases) %>%
  group_by(country, type) %>%
  summarise(total_cases = sum(cases)) %>%
  pivot_wider(names_from = type,
              values_from = total_cases)

confirmed_cases <- arrange(select(newest, country, confirmed), -confirmed)
deaths <- arrange(select(newest, country, death), -death)
recovered_cases <- arrange(select(newest, country, recovered), -recovered)
```

\clearpage
## Confirmed New Cases

These numbers are only as good as the amount of testing done and should be taken with a grain of salt.

Also keep in mind that there is a significant delay between being infected and having the results of the test.
Testing in most countries will only be done once symptoms appear and the test itself can also take a few days.
The numbers basically reflect the state of a few days in the past (the exact number of days varies from country to country).

```{r echo=FALSE}
head(confirmed_cases, 10)
```

```{r echo=FALSE}
ggplot(head(confirmed_cases,10), aes(reorder(country, confirmed), confirmed)) + 
  geom_bar(stat="identity", fill="black") +
  theme_economist() + 
  xlab("") + 
  theme(axis.text.x = element_text(angle = 90))
```


\clearpage
## New Deaths due to COVID-19

```{r echo=FALSE}
head(deaths, 10)
```

```{r echo=FALSE}
ggplot(head(deaths,10), aes(reorder(country, death), death)) + 
  geom_bar(stat="identity", fill="black") +
  theme_economist() + 
  xlab("") + 
  ylab("deaths") +
  theme(axis.text.x = element_text(angle = 90))
```

\clearpage
## New Recovered Cases

```{r echo=FALSE, position="left"}
head(recovered_cases, 10)
```


```{r echo=FALSE}
ggplot(head(recovered_cases,10), aes(reorder(country, recovered), recovered)) +
  geom_bar(stat="identity", fill="black") +
  theme_economist() + 
  xlab("") + 
  theme(axis.text.x = element_text(angle = 90))
```

\clearpage

# Trends

```{r echo=FALSE}
cases_df <- coronavirus %>% 
  group_by(Country.Region) %>% 
  arrange(date) %>% 
  filter(cases > 0)
cases_confirmed <- cases_df %>% filter(type == "confirmed")
cases_deaths <- cases_df %>% filter(type == "death")
cases_recovered <- cases_df %>% filter(type == "recovered")
```

```{r echo=FALSE, fig.height=2.6}
plot_trend <- function(cases_df, country, title) {
  cases_by_country <- cases_df %>% filter(Country.Region == country)
  ggplot(select(cases_by_country, date, cases, Country.Region), aes(date, cases)) + 
    geom_bar(stat="identity", fill="black") +
    theme_economist() +
    ggtitle(title) +
    xlab("") +
    ylab("")
}
```

## Worldwide

```{r echo=FALSE, fig.height=2.6}
ggplot(select(cases_confirmed,date,cases, Country.Region), aes(date, cases)) + 
  geom_bar(stat="identity", fill="black") +
  theme_economist() +
  ggtitle("New Confirmed Cases") +
  xlab("") +
  ylab("")
```

```{r echo=FALSE, fig.height=2.6}
ggplot(select(cases_recovered,date,cases, Country.Region), aes(date, cases)) + 
  geom_bar(stat="identity", fill="black") +
  theme_economist() +
  ggtitle("New Recovered Cases") +
  xlab("") +
  ylab("")
```

```{r echo=FALSE, fig.height=2.6}
ggplot(select(cases_deaths,date,cases, Country.Region), aes(date, cases)) + 
  geom_bar(stat="identity", fill="black") +
  theme_economist() +
  ggtitle("New Deaths due to COVID-19") +
  xlab("") +
  ylab("")
```

\clearpage

## Germany

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_confirmed, "Germany", "New Confirmed Cases")
```


```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_recovered, "Germany", "New Recovered Cases")
```

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_deaths, "Germany", "New Deaths due to COVID-19")
```
\clearpage

## France

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_confirmed, "France", "New Confirmed Cases")
```

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_recovered, "France", "New Recovered Cases")
```


```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_deaths, "France", "New Deaths due to COVID-19")
```

\clearpage

## Italy

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_confirmed, "Italy", "New Confirmed Cases")
```

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_recovered, "Italy", "New Recovered Cases")
```

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_deaths, "Italy", "New Deaths due to COVID-19")
```

\clearpage

## Spain

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_confirmed, "Spain", "New Confirmed Cases")
```

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_recovered, "Spain", "New Recovered Cases")
```

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_deaths, "Spain", "New Deaths due to COVID-19")
```

## United Kingdom

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_confirmed, "United Kingdom", "New Confirmed Cases")
```

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_recovered, "United Kingdom", "New Recovered Cases")
```

```{r echo=FALSE, fig.height=2.6}
plot_trend(cases_deaths, "United Kingdom", "New Deaths due to COVID-19")
```